---
title: "3.5 - Spatiotemporal Trends of Community Composition Across GRUMP Samples"
author: "Nathan WIlliams"
date: "2024-11-21"
output: html_document
---

```{r Load Dependencies}
library(tidyverse)
library(patchwork)
library(knitr)
library(ggpubr)
library(vegan)
library(pairwiseAdonis)
library(kableExtra)
```

```{r Import Data}
setwd("~/Dropbox/USC/Projects/grump.data/04-Data/version-1.3/")
grump_asv_long <- read_csv("grump_asv_long_version-1.3.1.csv")
```

```{r Grump Taxonomy}
grump_taxonomy <- grump_asv_long %>% select(Domain, Supergroup, Division, Phylum, Class, Order, Family, Genus, Species, ProPortal_ASV_Ecotype, Sequence_Type, ASV_hash, ASV) %>% distinct(ASV, .keep_all = TRUE)
```

```{r, echo=FALSE, fig1, fig.height = 10, fig.width = 15}
# Plot 1: Pie Chart of reads by Ocean Basin
pie.database <- grump_asv_long %>%
  group_by(Ocean.Basin) %>%
  summarise(Total.Reads = sum(Corrected_Sequence_Counts), .groups = 'drop') %>%
  mutate(Total = sum(Total.Reads),
         Total.Percent = (Total.Reads / Total) * 100)

pie.database$Total.Percent <- round(pie.database$Total.Percent, 1)

pie_chart <- ggplot(pie.database, aes(x = "", y = Total.Percent, fill = Ocean.Basin)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y") +
  theme_void() +
  geom_text(aes(label = paste0(Total.Percent, "%")), position = position_stack(vjust = 0.5)) +
  labs(title = "A) Total Reads by Ocean Basin") +
  scale_fill_brewer(palette = "Set3") +
  theme(legend.position = "bottom", legend.direction = "vertical") +
  guides(fill = guide_legend(title = "Ocean Basin", title.position = "top", title.hjust = 0.5))

# Plot 2: Pie Chart of reads by Depth
pie.database.depth <- grump_asv_long %>%
  group_by(Depth.Category) %>%
  summarise(Total.Reads = sum(Corrected_Sequence_Counts), .groups = 'drop') %>%
  mutate(Total = sum(Total.Reads),
         Total.Percent = (Total.Reads / Total) * 100)

pie.database.depth$Total.Percent <- round(pie.database.depth$Total.Percent, 1)

pie_chart_depth <- ggplot(pie.database.depth, aes(x = "", y = Total.Percent, fill = Depth.Category)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y") +
  theme_void() +
  geom_text(aes(label = paste0(Total.Percent, "%")), position = position_stack(vjust = 0.5)) +
  labs(title = "B) Total Reads by Depth") +
  scale_fill_brewer(palette = "Set3") +
  theme(legend.position = "bottom", legend.direction = "vertical") +
  guides(fill = guide_legend(title = "Depth Category", title.position = "top", title.hjust = 0.5))

# Plot 3: Pie Chart of samples by Ocean Basin
pie.samples <- grump_asv_long %>%
  group_by(Ocean.Basin) %>%
  summarise(Total.Samples = n_distinct(SampleID), .groups = 'drop') %>%
  mutate(Total = sum(Total.Samples),
         Total.Percent = (Total.Samples / Total) * 100)

pie.samples$Total.Percent <- round(pie.samples$Total.Percent, 1)

pie_chart_samples <- ggplot(pie.samples, aes(x = "", y = Total.Percent, fill = Ocean.Basin)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y") +
  theme_void() +
  geom_text(aes(label = paste0(Total.Percent, "%")), position = position_stack(vjust = 0.5)) +
  labs(title = "C) Total Samples by Ocean Basin") +
  scale_fill_brewer(palette = "Set3") +
  theme(legend.position = "bottom", legend.direction = "vertical") +
  guides(fill = guide_legend(title = "Ocean Basin", title.position = "top", title.hjust = 0.5))

# Plot 4: Pie Chart of samples by Depth
pie.samples.depth <- grump_asv_long %>%
  group_by(Depth.Category) %>%
  summarise(Total.Samples = n_distinct(SampleID), .groups = 'drop') %>%
  mutate(Total = sum(Total.Samples),
         Total.Percent = (Total.Samples / Total) * 100)

pie.samples.depth$Total.Percent <- round(pie.samples.depth$Total.Percent, 1)

pie_chart_samples_depth <- ggplot(pie.samples.depth, aes(x = "", y = Total.Percent, fill = Depth.Category)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y") +
  theme_void() +
  geom_text(aes(label = paste0(Total.Percent, "%")), position = position_stack(vjust = 0.5)) +
  labs(title = "D) Total Samples by Depth") +
  scale_fill_brewer(palette = "Set3") +
  theme(legend.position = "bottom", legend.direction = "vertical") +
  guides(fill = guide_legend(title = "Depth Category", title.position = "top", title.hjust = 0.5))

# Combine the plots in a 2x2 grid with the legend on the right
combined_plot <- (pie_chart | pie_chart_depth | plot_spacer()) / (pie_chart_samples | pie_chart_samples_depth | plot_spacer()) +
  plot_layout(guides = 'collect') & theme(legend.position = 'right', legend.box = "vertical")

# Print the combined plot
print(combined_plot)

# Print the tables
cat("Table 1: Total Reads and Percentages by Ocean Basin\n")
print(pie.database)

cat("Table 2: Total Reads and Percentages by Depth Category\n")
print(pie.database.depth)

cat("Table 3: Total Samples and Percentages by Ocean Basin\n")
print(pie.samples)

cat("Table 4: Total Samples and Percentages by Depth Category\n")
print(pie.samples.depth)

# Set directory and save the combined plot
#setwd("~/Dropbox/USC/Projects/grump.data/02-Figures")
#ggsave("Combined_GRUMP_Pie_Charts_with_Samples_Legend_Right.pdf", plot = combined_plot, width = 16, height = 8)
```

```{r Number of ASVs that exist in only one sample}
# Group by ASV and count the number of unique samples for each ASV
asv_sample_count <- grump_asv_long %>%
  group_by(ASV) %>%
  summarise(Sample_Count = n_distinct(SampleID))

# Filter ASVs that occur in only one sample
unique_asvs_in_one_sample <- asv_sample_count %>%
  filter(Sample_Count == 1)

# Calculate the total number of unique ASVs
total_unique_asvs <- nrow(asv_sample_count)

# Calculate the percentage of ASVs found in only one sample
percentage_single_sample_asvs <- (nrow(unique_asvs_in_one_sample) / total_unique_asvs) * 100

# Print the results
print(paste("Number of ASVs that exist in only one sample:", nrow(unique_asvs_in_one_sample)))
print(paste("Total number of unique ASVs:", total_unique_asvs))
print(paste("Percentage of ASVs that exist in only one sample:", round(percentage_single_sample_asvs, 2), "%"))
```

```{r Percentage of ASVs that exist in all samples}
# Group by ASV and count the number of unique samples for each ASV
asv_sample_count <- grump_asv_long %>%
  group_by(ASV) %>%
  summarise(Sample_Count = n_distinct(SampleID))

# Determine the total number of unique samples
total_samples <- n_distinct(grump_asv_long$SampleID)

# Filter ASVs that appear in all samples
asvs_in_all_samples <- asv_sample_count %>%
  filter(Sample_Count == total_samples)

# Calculate the percentage of ASVs found in all samples
percentage_all_samples_asvs <- (nrow(asvs_in_all_samples) / nrow(asv_sample_count)) * 100

# Print the results
print(paste("Number of ASVs that exist in all samples:", nrow(asvs_in_all_samples)))
print(paste("Total number of unique ASVs:", nrow(asv_sample_count)))
print(paste("Percentage of ASVs that exist in all samples:", round(percentage_all_samples_asvs, 2), "%"))
print("Here are some examples of ASVs found in all samples:")
```

```{r Number of ASVs that are detected within all Ocean Basin}
# Assuming 'Ocean.Basin' is a column in your dataframe that denotes the basin for each sample

# Group by ASV and Ocean.Basin, then count unique Ocean Basins per ASV
asv_basin_count <- grump_asv_long %>%
  group_by(ASV) %>%
  summarise(Ocean_Basins = n_distinct(Ocean.Basin))

# Determine the total number of unique Ocean Basins
total_ocean_basins <- n_distinct(grump_asv_long$Ocean.Basin)

# Filter ASVs that appear in all Ocean Basins
asvs_in_all_basins <- asv_basin_count %>%
  filter(Ocean_Basins == total_ocean_basins)

# Calculate the percentage of ASVs found in all Ocean Basins
percentage_all_basins_asvs <- (nrow(asvs_in_all_basins) / nrow(asv_basin_count)) * 100

# Print the results
print(paste("Number of ASVs present in every Ocean Basin:", nrow(asvs_in_all_basins)))
print(paste("Percentage of ASVs present in every Ocean Basin:", round(percentage_all_basins_asvs, 2), "%"))
```

```{r Top ASVs which are detected in all five oceans}
# Group by ASV and Ocean.Basin, then count unique Ocean Basins per ASV
asv_basin_count <- grump_asv_long %>%
  group_by(ASV) %>%
  summarise(Ocean_Basins = n_distinct(Ocean.Basin), .groups = 'drop')

# Determine the total number of unique Ocean Basins
total_ocean_basins <- n_distinct(grump_asv_long$Ocean.Basin)

# Filter ASVs that appear in all Ocean Basins
asvs_in_all_basins <- asv_basin_count %>%
  filter(Ocean_Basins == total_ocean_basins) %>%
  pull(ASV)  # Get a vector of ASVs

# Calculate the total abundance of each ASV across all samples, only for those ASVs in all basins
asv_abundance_totals <- grump_asv_long %>%
  filter(ASV %in% asvs_in_all_basins) %>%
  group_by(ASV) %>%
  summarise(Total_Abundance = sum(Relative_Abundance), .groups = 'drop') %>%
  arrange(desc(Total_Abundance))

# Get the top 10 most abundant ASVs from this filtered list
top_10_asvs_in_all_basins <- head(asv_abundance_totals, 10)

# Optionally, join back to grump_taxonomy to retrieve taxonomic information for these top 10 ASVs
top_10_asvs_with_taxonomy <- grump_taxonomy %>%
  filter(ASV %in% top_10_asvs_in_all_basins$ASV) %>%
  arrange(match(ASV, top_10_asvs_in_all_basins$ASV))  # Preserve order of abundance
```

```{r message = FALSE, warning = FALSE}
# Print the results
print("Top 10 most abundant ASVs that are present in all five oceans with their taxonomy:") 
 # Print the result for this specific Ocean Basin
  kbl(top_10_asvs_with_taxonomy) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>% 
  kable_paper() %>%
  scroll_box(width = "100%", height = "800px")

```

```{r ASVs that are unique to each single Ocean Basin}
library(dplyr)

# Assuming your dataframe is named grump_asv_long and has a column named 'Ocean.Basin'

# Group by ASV and count unique samples per ASV per Ocean Basin
asv_per_basin <- grump_asv_long %>%
  group_by(ASV, Ocean.Basin) %>%
  summarise(Sample_Count = n_distinct(SampleID), .groups = "drop")

# Self join to find ASVs that are only present in one Ocean Basin
asvs_unique_to_one_basin <- asv_per_basin %>%
  group_by(ASV) %>%
  filter(n() == 1) %>%
  ungroup() %>%
  distinct(ASV, Ocean.Basin) %>%
  arrange(Ocean.Basin, ASV)

# Calculate total number of unique ASVs
total_unique_asvs <- n_distinct(grump_asv_long$ASV)

# Print results by each Ocean Basin
unique_basins <- unique(asvs_unique_to_one_basin$Ocean.Basin)
print("Percentage of ASVs unique to each Ocean Basin relative to total unique ASVs:")
for(basin in unique_basins) {
  specific_basin_asvs <- filter(asvs_unique_to_one_basin, Ocean.Basin == basin)
  num_unique_basin_asvs <- nrow(specific_basin_asvs)
  percentage_basin_asvs <- (num_unique_basin_asvs / total_unique_asvs) * 100
  print(paste("Ocean Basin:", basin))
  print(paste("Number of unique ASVs:", num_unique_basin_asvs))
  print(paste("Percentage of unique ASVs in this basin:", round(percentage_basin_asvs, 2), "%"))
  print("Examples of unique ASVs:")
}
```

```{r The most abundant unique ASVs}
# Assuming grump_asv_long has columns 'ASV', 'Ocean.Basin', and 'Relative_Abundance'
# First, calculate the total relative abundance for each ASV within each Ocean Basin
asv_abundance_per_basin <- grump_asv_long %>%
  group_by(Ocean.Basin, ASV) %>%
  summarise(Total_Abundance = sum(Relative_Abundance), .groups = 'drop')

# Now, for each Ocean Basin, rank the ASVs by their total relative abundance and select the top 10
unique_oceans <- unique(asv_abundance_per_basin$Ocean.Basin)

for (ocean in unique_oceans) {
  top_asvs <- asv_abundance_per_basin %>%
    filter(Ocean.Basin == ocean) %>%
    arrange(desc(Total_Abundance)) %>%
    slice_max(order_by = Total_Abundance, n = 10)
  
  # Optionally, include taxonomic information if available
  if (exists("grump_taxonomy")) {
    top_asvs <- top_asvs %>%
      left_join(grump_taxonomy, by = "ASV")
  }}
```

```{r message = FALSE, warning = FALSE}
#Print results
  kbl(top_asvs) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>% 
  kable_paper() %>%
  scroll_box(width = "100%", height = "800px")
```

```{r }
#Make Taxonomy table
grump_taxonomy <- grump_asv_long %>% select(Domain, Supergroup, Division, Phylum, Class, Order, Family, Genus, Species, ProPortal_ASV_Ecotype, Sequence_Type, ASV_hash, ASV) %>% distinct(ASV, .keep_all = TRUE)
```

```{r }
# Define custom colors for Longhurst provinces
longhurst_colours <- c("Polar - Boreal Polar Province (POLR)"="#fe0080", 
                       "Polar - N. Pacific Epicontinental Province"="#ffcbd7", 
                       "Westerlies - Pacific Subarctic Gyres Province (East)"="#6e000b", 
                       "Coastal - NW Atlantic Shelves Province"="#f75a00", 
                       "Westerlies - Gulf Stream Province"="#734900", 
                       "Westerlies - N. Atlantic Subtropical Gyral Province (East) (STGE)"="#ffc274", 
                       "Westerlies - N. Pacific Polar Front Province"="#ffe1a4", 
                       "Westerlies - N. Atlantic Subtropical Gyral Province (West) (STGW)"="#b49200", 
                       "Trades - N. Pacific Tropical Gyre Province"="#cebd00", 
                       "Trades - N. Atlantic Tropical Gyral Province (TRPG)"="#688900", 
                       "Coastal - E. India Coastal Province"="#8ffd32", 
                       "Trades - Western Tropical Atlantic Province"="#cfffa6", 
                       "Trades - N. Pacific Equatorial Countercurrent Province"="#0d1c00", 
                       "Trades - Pacific Equatorial Divergence Province"="#005d07", 
                       "Trades - South Atlantic Gyral Province (SATG)"="#7fffae", 
                       "Westerlies - S. Pacific Subtropical Gyre Province"="#00af64", 
                       "Trades - Indian Monsoon Gyres Province"="#007859", 
                       "Trades - Archipelagic Deep Basins Province"="#c5fcff", 
                       "Coastal - East Australian Coastal Province"="#004a4f", 
                       "Coastal - E. Africa Coastal Province"="#01b0be", 
                       "Coastal - Benguela Current Coastal Province"="#0058b3", 
                       "Trades - Indian S. Subtropical Gyre Province"="#b8aeff", 
                       "Westerlies - S. Subtropical Convergence Province"="#21003f", 
                       "Coastal - New Zealand Coastal Province"="#eac7ff", 
                       "Coastal - SW Atlantic Shelves Province"="#7b00a5", 
                       "Westerlies - Subantarctic Province"="#d95dff", 
                       "Polar - Antarctic Province"="#630037", 
                       "Polar - Austral Polar Province"="#ff8fb9")

ordered_longhurst_provinces <- rev(c("Polar - Boreal Polar Province (POLR)", "Polar - N. Pacific Epicontinental Province", "Westerlies - Pacific Subarctic Gyres Province (East)", "Coastal - NW Atlantic Shelves Province", "Westerlies - Gulf Stream Province", "Westerlies - N. Atlantic Subtropical Gyral Province (East) (STGE)", "Westerlies - N. Pacific Polar Front Province", "Westerlies - N. Atlantic Subtropical Gyral Province (West) (STGW)", "Trades - N. Pacific Tropical Gyre Province", "Trades - N. Atlantic Tropical Gyral Province (TRPG)", "Coastal - E. India Coastal Province", "Trades - Western Tropical Atlantic Province", "Trades - N. Pacific Equatorial Countercurrent Province", "Trades - Pacific Equatorial Divergence Province", "Trades - South Atlantic Gyral Province (SATG)", "Westerlies - S. Pacific Subtropical Gyre Province", "Trades - Indian Monsoon Gyres Province", "Trades - Archipelagic Deep Basins Province", "Coastal - East Australian Coastal Province", "Coastal - E. Africa Coastal Province", "Coastal - Benguela Current Coastal Province", "Trades - Indian S. Subtropical Gyre Province", "Westerlies - S. Subtropical Convergence Province", "Coastal - New Zealand Coastal Province",
"Coastal - SW Atlantic Shelves Province", "Westerlies - Subantarctic Province", "Polar - Antarctic Province", "Polar - Austral Polar Province"))
```

```{r include = FALSE, message = FALSE, warning = FALSE}
#Pull out the top 50m
Top50m_grump <- grump_asv_long %>% filter(Depth < 50)
```

Average by grouping by station, Cruise_ID, and Level_1.1. For this script Level_1.1 will be abbreviated to Level_1.1.

Average by grouping by station, Cruise_ID, and Level_1.1.

TS = Total Species at site
TP = total sum of the populations
Ara = Average Relative_Abundance of Level_1.1 at this particular pre-defined depth

```{r include = FALSE, message = FALSE, warning = FALSE}
#Calculate Relative_Abundance for the Level_1.1 category
Level_1.1.Surface      <- Top50m_grump %>% select("Cruise_ID", "Station", "Date", "Latitude", "Longitude","ASV_hash", "Level_1.1","Corrected_Sequence_Counts")
Level_1.1.Surface      <- Level_1.1.Surface   %>% group_by(Cruise_ID,Station,Level_1.1,Date) %>% mutate(ASC = mean(Corrected_Sequence_Counts))
Level_1.1.Surface      <- Level_1.1.Surface   %>% group_by(Cruise_ID,Station,Level_1.1,Date) %>% mutate(TS = sum(ASC))
Level_1.1.Surface      <- Level_1.1.Surface   %>% group_by(Cruise_ID,Station,Date) %>% mutate(TP = sum(ASC))
Level_1.1.Surface      <- Level_1.1.Surface   %>% group_by(Cruise_ID,Station,Date,Level_1.1) %>% mutate(Level_1.1.Surface_ra = (TS/TP)*100)
Level_1.1.Surface      <- Level_1.1.Surface   %>% unite("Distinct.var",Cruise_ID,Station,Date,Level_1.1, remove = FALSE) %>% distinct(Distinct.var, .keep_all = TRUE)
#Level_1.1.Surface[is.na(Level_1.1.Surface)] <- 0
Level_1.1.Surface      <- Level_1.1.Surface   %>% group_by(Cruise_ID,Station,Date) %>% mutate(Check.Total = sum(Level_1.1.Surface_ra))
```

Surface only zoomed in for 3 major transects.

Make the three bar graphs.

Make the colours and the order of sequence type
```{r include = FALSE, message = FALSE, warning = FALSE}
#Seqeuence Type colour and order
bar_width <- 0.75
text_size <- 18
axis_text_angle <- 90

# Colours and order for Level_1.1
Level_1.1.order <- c("Archaea", "Bacteria", "Cyanobacteria", "Chloroplast_16S", "Eukaryote_18S")
Level_1.1.colours <- c("Archaea" = "#e32d24", "Eukaryote_18S" = "#009b80", "Bacteria" = "#aaa4d8", "Chloroplast_16S" = "#a8cb40", "Cyanobacteria" = "#ffc544")
```

```{r include = FALSE, message = FALSE, warning = FALSE}
#Functions

# Function to calculate relative abundance for a given depth category
calculate_relative_abundance_depth <- function(data, depth_category) {
  data %>% 
    filter(Depth.Category %in% depth_category) %>% 
    select("Cruise_ID", "Station", "Date", "Latitude", "Longitude", "ASV_hash", "Level_1.1", "Corrected_Sequence_Counts","Ocean.Basin", "Longhurst_Long","Depth","Depth.Category") %>% 
    group_by(Cruise_ID, Station, Level_1.1, Date) %>% 
    mutate(ASC = mean(Corrected_Sequence_Counts),
           TS = sum(ASC)) %>% 
    group_by(Cruise_ID, Station, Date) %>% 
    mutate(TP = sum(ASC)) %>% 
    group_by(Cruise_ID, Station, Date, Level_1.1) %>% 
    mutate(Level_1.1.ra = (TS / TP) * 100) %>% 
    unite("Distinct.var", Cruise_ID, Station, Date, Level_1.1, remove = FALSE) %>% 
    distinct(Distinct.var, .keep_all = TRUE) %>% 
    group_by(Cruise_ID, Station, Date) %>% 
    mutate(Check.Total = sum(Level_1.1.ra))
}

# Function to calculate relative abundance for a given depth category
calculate_relative_abundance <- function(data) {
  data %>% 
    select("Cruise_ID", "Station", "Date", "Latitude", "Longitude", "ASV_hash", "Level_1.1", "Corrected_Sequence_Counts","Ocean.Basin", "Longhurst_Long","Depth", "Depth.Category") %>% 
    group_by(Cruise_ID, Station, Level_1.1, Date) %>% 
    mutate(ASC = mean(Corrected_Sequence_Counts),
           TS = sum(ASC)) %>% 
    group_by(Cruise_ID, Station, Date) %>% 
    mutate(TP = sum(ASC)) %>% 
    group_by(Cruise_ID, Station, Date, Level_1.1) %>% 
    mutate(Level_1.1.ra = (TS / TP) * 100) %>% 
    unite("Distinct.var", Cruise_ID, Station, Date, Level_1.1, remove = FALSE) %>% 
    distinct(Distinct.var, .keep_all = TRUE) %>% 
    group_by(Cruise_ID, Station, Date) %>% 
    mutate(Check.Total = sum(Level_1.1.ra))
}

# Function to plot bar graph
plot_bar_graph <- function(data, y_label, limits, breaks) {
  ggplot(data, aes(x = Latitude, y = Level_1.1.ra, fill = factor(Level_1.1, levels = Level_1.1.order))) + 
    geom_bar(stat = 'identity', colour = "black", size = 0.25, width = bar_width) +
    theme(legend.position = 'right') + 
    scale_fill_manual(values = Level_1.1.colours, "Level_1.1") +
    guides(fill = guide_legend(ncol = 1)) + 
    labs(y = y_label, x = "") +
    scale_x_continuous(limits = limits, breaks = breaks) +
    theme_minimal(base_size = 15) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 20)) + 
    theme(axis.text.y = element_text(size = 20)) + 
    coord_flip()
}

plot_bar_graph_horizontal <- function(data, x_label, limits, breaks) {
  # Filter out rows where Level_1.1.ra is zero or missing
  data_filtered <- data %>% filter(Level_1.1.ra != 0 & !is.na(Level_1.1.ra))
  
  ggplot(data_filtered, aes(y = Level_1.1.ra, x = Latitude, fill = factor(Level_1.1, levels = Level_1.1.order))) + 
    geom_bar(stat = 'identity', colour = "black", size = 0.25, width = bar_width) +
    theme(legend.position = 'right') + 
    scale_fill_manual(values = Level_1.1.colours, "Level_1.1") + 
    guides(fill = guide_legend(ncol = 1)) +
    labs(y = x_label, x = "") +  # Notice that x and y labels are flipped
    scale_x_continuous(limits = limits, breaks = breaks) +  # Notice that limits and breaks apply to x-axis now
    theme_minimal(base_size = 15) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 20)) + 
    theme(axis.text.y = element_text(size = 20))
}

plot_bar_graph_MOSAiC <- function(data, x_label, limits, breaks) {
  # Filter out rows where Level_1.1.ra is zero or missing
  data_filtered <- data %>% filter(Level_1.1.ra != 0 & !is.na(Level_1.1.ra))
  
  ggplot(data_filtered, aes(y = Level_1.1.ra, x = Longitude, fill = factor(Level_1.1, levels = Level_1.1.order))) + 
    geom_bar(stat = 'identity', colour = "black", size = 0.25, width = bar_width) +
    theme(legend.position = 'right') + 
    scale_fill_manual(values = Level_1.1.colours, "Level_1.1") + 
    guides(fill = guide_legend(ncol = 1)) +
    labs(y = x_label, x = "") +  # Notice that x and y labels are flipped
    scale_x_continuous(limits = limits, breaks = breaks) +  # Notice that limits and breaks apply to x-axis now
    theme_minimal(base_size = 15) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 20)) + 
    theme(axis.text.y = element_text(size = 20))
}

calculate_summary_stats <- function(data) {
  data %>%
    group_by(Level_1.1, Ocean.Basin) %>%
    summarise(
      count = n(),
      mean = round(mean(Level_1.1.ra, na.rm = TRUE), 2),
      sd = round(sd(Level_1.1.ra, na.rm = TRUE), 2),
      min = round(min(Level_1.1.ra, na.rm = TRUE), 2),
      `25%` = round(quantile(Level_1.1.ra, 0.25, na.rm = TRUE), 2),
      median = round(median(Level_1.1.ra, na.rm = TRUE), 2),
      `75%` = round(quantile(Level_1.1.ra, 0.75, na.rm = TRUE), 2),
      max = round(max(Level_1.1.ra, na.rm = TRUE), 2)
    )
}

calculate_summary_stats_depths <- function(data) {
  data %>%
    group_by(Level_1.1, Ocean.Basin, Depth.Category) %>%
    summarise(
      count = n(),
      mean = round(mean(Level_1.1.ra, na.rm = TRUE), 2),
      sd = round(sd(Level_1.1.ra, na.rm = TRUE), 2),
      min = round(min(Level_1.1.ra, na.rm = TRUE), 2),
      `25%` = round(quantile(Level_1.1.ra, 0.25, na.rm = TRUE), 2),
      median = round(median(Level_1.1.ra, na.rm = TRUE), 2),
      `75%` = round(quantile(Level_1.1.ra, 0.75, na.rm = TRUE), 2),
      max = round(max(Level_1.1.ra, na.rm = TRUE), 2)
    )
}
```

```{r message = FALSE, warning = FALSE}
#Define Top 50 and Calculate relative abundance
Top50m_grump <- grump_asv_long %>% filter(Depth < 50)
Level_1.1.Surface <- calculate_relative_abundance_depth(Top50m_grump, "Depth_0m-200m")
```

```{r message = FALSE, warning = FALSE}
#Summary Statistics
Level_1.1.Surface.summary_stats <- calculate_summary_stats(Level_1.1.Surface)
kbl(Level_1.1.Surface.summary_stats) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>% 
  kable_paper() %>%
  scroll_box(width = "100%", height = "800px")
```

```{r message = FALSE, warning = FALSE}
# Pairwise Adonis in top 50m
Level_1.1.Surface_data <- Level_1.1.Surface %>% rename(Relative_Abundance = Level_1.1.ra) %>% unite("Pairwise.Var", Level_1.1, Ocean.Basin, sep = "_in_")
# Ensure that the required columns are in the correct format
Level_1.1.Surface_data$Level_1.1 <- as.factor(Level_1.1.Surface_data$Pairwise.Var)
Level_1.1.Surface_data$Relative_Abundance <- as.numeric(as.character(Level_1.1.Surface_data$Relative_Abundance))

# Remove rows with NA in the Relative_Abundance, Level_1.1, or Ocean.Basin columns
Level_1.1.Surface_data <- na.omit(Level_1.1.Surface_data[, c("Relative_Abundance", "Pairwise.Var")])

# Create a distance matrix for the Relative_Abundance
dist_matrix <- vegdist(Level_1.1.Surface_data$Relative_Abundance, method = "bray")

# Run pairwise adonis
pairwise_results <- pairwise.adonis(dist_matrix, Level_1.1.Surface_data$Pairwise.Var)

# Print results
kbl(pairwise_results) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>% 
  kable_paper() %>%
  scroll_box(width = "100%", height = "800px")
```

```{r Generate plots, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 8}
Level_1.1.GA02 <- Level_1.1.Surface %>% filter(Cruise_ID %in% c("GA02","GA03","AMT-19","AMT-20")) %>% filter(!Station == "Station_9")
Level_1.1.P16SP16N <- Level_1.1.Surface %>% filter(Cruise_ID %in% c("P16S","P16N","POTATOE"))
Level_1.1.I8S_I9N <- Level_1.1.Surface %>% filter(Cruise_ID %in% c("I8S", "I9N"))

Level_1.1.GA02_sbg <- plot_bar_graph(Level_1.1.GA02, "Atlantic Ocean - GA02 GA03 AMT19 AMT20", c(-45, 45), seq(-45, 45, by = 10))
Level_1.1.P16SP16N_sbg <- plot_bar_graph(Level_1.1.P16SP16N, "Pacific Ocean - P16N P16S P15S POTATOE", c(-45, 45), seq(-45, 45, by = 10))
Level_1.1.I8S_I9N_sbg <- plot_bar_graph(Level_1.1.I8S_I9N, "Indian Ocean - IO8S IO9N", c(-45, 45), seq(-45, 45, by = 10))

Level_1.1.MOSAiC <- Level_1.1.Surface %>% filter(Cruise_ID %in% c("MOSAiC"))
Level_1.1.MOSAiC_sbg <- plot_bar_graph_MOSAiC(Level_1.1.MOSAiC, "MOSAiC", c(-10, 120), seq(-10, 120, by = 10))
Level_1.1.MOSAiC_sbg

Top_50m_Three_Oceans <- ggarrange(Level_1.1.P16SP16N_sbg, Level_1.1.GA02_sbg, Level_1.1.I8S_I9N_sbg, ncol = 3, nrow = 1, common.legend = TRUE, legend = "right")
Top_50m_Three_Oceans

Southern_Ocean_P15S <- Level_1.1.Surface %>% filter(Cruise_ID %in% c("P15S"))
Southern_Ocean_I8S <- Level_1.1.Surface %>% filter(Cruise_ID %in% c("I8S"))
Southern_Ocean_IND_2017 <- Level_1.1.Surface %>% filter(Cruise_ID %in% c("IND_2017"))

Southern_Ocean_P15S_sbg <- plot_bar_graph(Southern_Ocean_P15S, "Southern Ocean P15S", c(-70, -45), seq(-70, -45, by = 10))
Southern_Ocean_I8S_sbg <- plot_bar_graph(Southern_Ocean_I8S, "Southern Ocean I8S", c(-70, -45), seq(-70, -45, by = 10))
Southern_Ocean_IND_2017_sbg <- plot_bar_graph(Southern_Ocean_IND_2017, "Southern Ocean IND 2017", c(-70, -45), seq(-70, -45, by = 10))

Southern_Ocean_P15S_sbg
Southern_Ocean_I8S_sbg
Southern_Ocean_IND_2017_sbg
```

```{r message = FALSE, warning = FALSE}
#Summary Statistics Depth 0-200m
Level_1.1.D0m_D200m <- calculate_relative_abundance_depth(grump_asv_long, "Depth_0m-200m")
Level_1.1.D200m_D1000m <- calculate_relative_abundance_depth(grump_asv_long, "Depth_200m-1000m")
Level_1.1.D_Below_1000m <- calculate_relative_abundance_depth(grump_asv_long, "Depth_>1000m")

#Make the stats tables
Level_1.1.D0m_D200m.summary_stats <- calculate_summary_stats_depths(Level_1.1.D0m_D200m)
Level_1.1.D200m_D1000m.summary_stats <- calculate_summary_stats_depths(Level_1.1.D200m_D1000m)
Level_1.1.D_Below_1000m.summary_stats <- calculate_summary_stats_depths(Level_1.1.D_Below_1000m)


# Combine the three tables
Level_1.1.Depth.summary_stats <- bind_rows(
  Level_1.1.D0m_D200m.summary_stats,
  Level_1.1.D200m_D1000m.summary_stats,
  Level_1.1.D_Below_1000m.summary_stats
)

# Display the combined table with kable
kbl(Level_1.1.Depth.summary_stats) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "800px")
```
  

```{r message = FALSE, warning = FALSE, fig.width = 12, fig.height = 8}
#Depth Plots for the Pacific Ocean
Level_1.1_P16NS_D0m_D200m <- Level_1.1.D0m_D200m %>% filter(Cruise_ID %in% c("P16N", "P16S"))
Level_1.1_P16NS_D200m_D1000 <- Level_1.1.D200m_D1000m %>% filter(Cruise_ID %in% c("P16N", "P16S"))
Level_1.1_P16NS_D_Below_1000m <- Level_1.1.D_Below_1000m %>% filter(Cruise_ID %in% c("P16N", "P16S"))

Level_1.1_P16NS_D0m_D200m_sbg <- plot_bar_graph_horizontal(Level_1.1_P16NS_D0m_D200m, "P16NS_D0m_D200m", c(-45, 45), seq(-45, 45, by = 10))
Level_1.1_P16NS_D200m_D1000_sbg <- plot_bar_graph_horizontal(Level_1.1_P16NS_D200m_D1000, "P16NS_D200m_D1000", c(-45, 45), seq(-45, 45, by = 10))
Level_1.1_P16NS_D_Below_1000m_sbg <- plot_bar_graph_horizontal(Level_1.1_P16NS_D_Below_1000m, "P16NS_D_Below_1000m", c(-45, 45), seq(-45, 45, by = 10))

Pacific_Ocean_Depth_Profile <- ggarrange(Level_1.1_P16NS_D0m_D200m_sbg, Level_1.1_P16NS_D200m_D1000_sbg, Level_1.1_P16NS_D_Below_1000m_sbg, ncol = 1, nrow = 3, common.legend = TRUE, legend = "right")
Pacific_Ocean_Depth_Profile
#ggsave("Pacific_Ocean_Depth_Profile.pdf", width = 12, height = 5.85)
```

```{r message = FALSE, warning = FALSE,  fig.width = 12, fig.height = 8}
#Depth Plots for the Indian Ocean
Level_1.1_I9N_I8S_D0m_D200m <- Level_1.1.D0m_D200m %>% filter(Cruise_ID %in% c("I9N", "I8S"))
Level_1.1_I9N_I8S_D200m_D1000 <- Level_1.1.D200m_D1000m %>% filter(Cruise_ID %in% c("I9N", "I8S"))
Level_1.1_I9N_I8S_D_Below_1000m <- Level_1.1.D_Below_1000m %>% filter(Cruise_ID %in% c("I9N", "I8S"))

Level_1.1_I9N_I8S_D0m_D200m_sbg <- plot_bar_graph_horizontal(Level_1.1_I9N_I8S_D0m_D200m, "I9N.I8S_D0m_D200m", c(-45, 45), seq(-45, 45, by = 10))
Level_1.1_I9N_I8S_D200m_D1000_sbg <- plot_bar_graph_horizontal(Level_1.1_I9N_I8S_D200m_D1000, "I9N.I8S_D200m_D1000", c(-45, 45), seq(-45, 45, by = 10))
Level_1.1_I9N_I8S_D_Below_1000m_sbg <- plot_bar_graph_horizontal(Level_1.1_I9N_I8S_D_Below_1000m, "I9N.I8S_D_Below.1000m", c(-45, 45), seq(-45, 45, by = 10))

Indian_Ocean_Depth_Profile <- ggarrange(Level_1.1_I9N_I8S_D0m_D200m_sbg, Level_1.1_I9N_I8S_D200m_D1000_sbg, Level_1.1_I9N_I8S_D_Below_1000m_sbg, ncol = 1, nrow = 3, common.legend = TRUE, legend = "right")
Indian_Ocean_Depth_Profile
#ggsave("Indian_Ocean_Depth_Profile.pdf", width = 12, height = 5.85)
```

```{r message = FALSE, warning = FALSE, fig.width = 12, fig.height = 8}
# Depth Plots for Atlantic Ocean
Level_1.1_GA02_D0m_D200m <- Level_1.1.D0m_D200m %>% filter(Cruise_ID == "GA02")
Level_1.1_GA02_D200m_D1000 <- Level_1.1.D200m_D1000m %>% filter(Cruise_ID == "GA02")
Level_1.1_GA02_D_Below_1000m <- Level_1.1.D_Below_1000m %>% filter(Cruise_ID == "GA02")

Level_1.1_GA02_D0m_D200m_sbg <- plot_bar_graph_horizontal(Level_1.1_GA02_D0m_D200m, "GA02_D0m_D200m", c(-45, 45), seq(-45, 45, by = 10))
Level_1.1_GA02_D200m_D1000_sbg <- plot_bar_graph_horizontal(Level_1.1_GA02_D200m_D1000, "GA02_D200m_D1000", c(-45, 45), seq(-45, 45, by = 10))
Level_1.1_GA02_D_Below_1000m_sbg <- plot_bar_graph_horizontal(Level_1.1_GA02_D_Below_1000m, "GA02_D_Below.1000m", c(-45, 45), seq(-45, 45, by = 10))

Atlantic_Ocean_Depth_Profile <- ggarrange(Level_1.1_GA02_D0m_D200m_sbg, Level_1.1_GA02_D200m_D1000_sbg, Level_1.1_GA02_D_Below_1000m_sbg, ncol = 1, nrow = 3, common.legend = TRUE, legend = "right")
Atlantic_Ocean_Depth_Profile
#ggsave("Atlantic_Ocean_Depth_Profile.pdf", width = 12, height = 5.85)
```

```{r message = FALSE, warning = FALSE, fig.width = 12, fig.height = 8}
#Depth Plots for Southern Ocean
Level_1.1_P15S_D0m_D200m <- Level_1.1.D0m_D200m %>% filter(Cruise_ID == "P15S") %>% filter(!Station == "Station_30")
Level_1.1_P15S_D200m_D1000 <- Level_1.1.D200m_D1000m %>% filter(Cruise_ID == "P15S")
Level_1.1_P15S_D_Below_1000m <- Level_1.1.D_Below_1000m %>% filter(Cruise_ID == "P15S") %>% filter(!Station %in% c("Station_12", "Station_11"))

Level_1.1_P15S_D0m_D200m_sbg <- plot_bar_graph_horizontal(Level_1.1_P15S_D0m_D200m, "P15S_D0m_D200m", c(-70, -45), seq(-70, -45, by = 5))
Level_1.1_P15S_D200m_D1000_sbg <- plot_bar_graph_horizontal(Level_1.1_P15S_D200m_D1000, "P15S_D200m_D1000", c(-70, -45), seq(-70, -45, by = 5))
Level_1.1_P15S_D_Below_1000m_sbg <- plot_bar_graph_horizontal(Level_1.1_P15S_D_Below_1000m, "P15S_D_Below.1000m", c(-70, -45), seq(-70, -45, by = 5))

Southern_Ocean_Depth_Profile <- ggarrange(Level_1.1_P15S_D0m_D200m_sbg, Level_1.1_P15S_D200m_D1000_sbg, Level_1.1_P15S_D_Below_1000m_sbg, ncol = 1, nrow = 3, common.legend = TRUE, legend = "right")
Southern_Ocean_Depth_Profile
#ggsave("Southern_Ocean_Depth_Profile.pdf", width = 12, height = 5.85)
```
